Northern Pikeminnow Abundance in Deadwater Slough, Salmon River, Idaho, and Potential Impacts to Local Chinook Salmon Populations

### 2022-04-01 analysis/paper/paper.Rmd
moved Avian Predation to Supplementary Materials
minor edits in Discussion prior to near final editing

## Avian Predation

@@

Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere (@Winther2020). A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues. Bounty programs could be conducted year-round or seasonally to coincide with peak juvenile outmigration(s). Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.

@@

The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow. Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkably high density given the size of the area. Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon. We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than 24,000 juvenile Chinook salmon, however, it is likely that this value is closer to 100,000, annually. The magnitude of predation by Northern Pikeminnow can reduce adult escapement between 47% and 97% and should be considered a major factor influencing the recovery ESA-listed Chinook Salmon populations in the Upper Salmon MPG.

@@

>>>>

@@

Northern Pikeminnow bounty programs have proven successful at reducing population sizes and impacts on emigrating salmonids elsewhere [@Winther2020]. A local Northern Pikeminnow bounty program could provide monetary incentive for sportfisherman, in addition to boosting the local economy from lost revenues. Bounty programs could be conducted year-round or seasonally to coincide with peak juvenile outmigration(s). Each approach would require continual/annual effort and harvest to suppress the population size and as such, appropriate funding would be needed to support the program.

@@

The failure of a mining reservoir dam in the late 19th century created an alluvial fan at the confluence of the Salmon River and Dump Creek which has backed up the Salmon River and created favorable conditions for Northern Pikeminnow. Even under the most conservative scenarios, we estimate more than 10,000 Northern Pikeminnow occupy Deadwater Slough, resulting in a remarkably high density given the size of the area. Additionally, the slow water velocity and lack of cover have created conditions that favor predation on juvenile salmonids that are rearing in or emigrating through Deadwater Slough, including Chinook Salmon. We estimate that, at a minimum, Northern Pikeminnow in Deadwater Slough consume greater than `r prettyNum(round(min(total_impacts$juv_Lci), -3), big.mark = ",")` juvenile Chinook salmon, however, it is likely that this value is closer to `r prettyNum(round(total_impacts$juv_N[total_impacts$perc_diet == 60 & total_impacts$p_chnk == 0.50], -3), big.mark = ",")`, annually. The magnitude of predation by Northern Pikeminnow can reduce adult escapement between `r round((min(total_impacts$adult_N) / 558) * 100)`% and `r round((max(total_impacts$adult_N) / 558) * 100)`% and should be considered a major factor influencing the recovery ESA-listed Chinook Salmon populations in the Upper Salmon MPG.

@@

# Supplementary Materials
